\sysfull

### 2018-08-03 paper/asplos19/background.tex
revice the background

\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and
future energy-harvesting platforms based around general purpose, commodity
computing components~\cite{wisp,msp430datasheet}. We assume a device with a
memory system that has fast, byte-addressable volatile and non-volatile memory;
in particular, our target platform, WISP~\cite{wisp}, is equipped with a
mixture of SRAM and FRAM. Our implementation leverages hardware support for
fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys do not
require architectural additions to commodity processors as
in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}. \sys supports I/O
behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and
unsafe, asynchronous I/O.

@@

Checkpointing volatile state alone do not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Especially, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca} ---see
Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
inconsistent in an intermittent execution using the cyclic redundancy check
(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
maintain data consistent in the presence of WAR dependency is necessary for
correct execution.
%Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}. Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value. The situation occurs when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
%Figure~\ref{fig:code_demo_incosistency} illgit statuustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}. The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
%Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5. Figure~\ref{fig:crc_example_war} shows an intermittent execution. The second iteration writes $\texttt{Rmnd}$, but power fails before the checkpoint. After restarting, line 9 reads the updated value of $\texttt{Rmnd}$, producing an incorrect $\texttt{data}$ value. Checkpoint-based systems risk violating memory consistency~\cite{dino}. 
%To maintain consistency, prior systems~\cite{dino} {\em version} a subset of non-volatile data with the checkpoint. Task-based systems~\cite{chain,alpaca}, which we focus on, ensure consistency using programmer, compiler, and runtime support.

@@

Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
decompose their program into tasks, which is a region of code that can contain
arbitrary computation, sensing, and communication.  Task-based models keep
progress at the granularity of tasks by repeatedly trying to re-execute each
task on power failure until it successfully finishes, only then moving on to
the next task.  Since these models do not rely on taking an expensive
checkpoint, it is usually faster than the checkpoint-based
solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
task-based programming model, where the programmer explicitly expresses the
application as a sequence of tasks and the transition between them.
%Task-based runtime systems ensure that repeated task executions are idempotent. 
%The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.  
%A programmer describes task control-flow as a {\em task graph}. Task flow happens at programmer demarcated transition points that may be conditional on program values. 
%Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.

@@

%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.  
% We do not need to emphasize the following figure! (Sinan)
%\begin{figure}
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
%	\caption{Cost of \emph{write} operation}
%\end{subfigure}%
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
%	\caption{Cost of \emph{read} operation}
%\end{subfigure}
%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
%\end{figure}
As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
voltage drop on a capacitor powering the device across 1600 read and write
accesses to each memory, randomizing accesses to avoid caching effects. Our
data showed that a non-volatile memory access consumes around 1.5 times the
energy of a volatile access. The data confirm that, because non-volatile memory
accesses are more expensive than volatile memory accesses, prior systems that
frequently access non-volatile memory may be artificially inefficient.
%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. This observation is also consistent with the TI specification~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek} 
\sys is motivated by this observation: \sys 
moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
accesses, to decrease energy consumption and increase the time between power
failures.

@@

\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or
programmer-defined tasks can both both incorrect and non-terminating or
inefficient.  If code between consecutive checkpoints or in a task (we refer to
both scenarios as ''code in a task'' in this section) consumes {\em more} than the fixed,
maximum energy that the device can buffer, then the task
will never be able to complete using buffered energy.  Such a task is
non-terminating, {\em prevents forward progress}, and makes the program
incorrect. 
%
If the task consumes far {\em less} energy than a device can buffer, the system
may operate {\em inefficiently}, checkpointing or transitioning between tasks
more often than needed. 
%
Avoiding excessively costly, non-terminating tasks and short, high-overhead
tasks is challenging both to the programmer and the compiler, because
estimating the exact energy use of an arbitrary code is complicated.  When {\em
porting} a program from one platform to another, the problem becomes worse
because an excessively costly task on one device may be a relatively short task
on a device with a larger buffer.  Buffer sizes on different platforms vary
widely from, e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
complicating the problem.
%
As a solution, \sys's execution model presents {\em task coalescing}, which
merges multiple tasks to amortize the overhead when tasks are too small, and
{\em task downscaling}, which breaks up a task when the task is too large to complete. 
%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity. Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer. The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead. \sys's execution model based on {\em task coalescing} and {\em task downscaling}.  Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately. Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.

>>>>

\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and
future energy-harvesting platforms based around general purpose, commodity
computing components~\cite{wisp,msp430datasheet}. We assume a device with a
memory system that has fast, byte-addressable volatile and non-volatile memory;
in particular, our target platform, WISP~\cite{wisp}, is equipped with a
mixture of SRAM and FRAM. Our implementation leverages hardware support for
fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys do not
require architectural additions to commodity processors as
in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

Checkpointing volatile state alone do not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Especially, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
inconsistent in an intermittent execution using the cyclic redundancy check
(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
maintain data consistent in the presence of WAR dependency is necessary for
correct execution.}
%Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}. Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value. The situation occurs when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
%Figure~\ref{fig:code_demo_incosistency} illgit statuustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}. The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
%Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5. Figure~\ref{fig:crc_example_war} shows an intermittent execution. The second iteration writes $\texttt{Rmnd}$, but power fails before the checkpoint. After restarting, line 9 reads the updated value of $\texttt{Rmnd}$, producing an incorrect $\texttt{data}$ value. Checkpoint-based systems risk violating memory consistency~\cite{dino}. 
%To maintain consistency, prior systems~\cite{dino} {\em version} a subset of non-volatile data with the checkpoint. Task-based systems~\cite{chain,alpaca}, which we focus on, ensure consistency using programmer, compiler, and runtime support.

@@

Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
decompose their program into \textbf{tasks}, which is a region of code that can contain
arbitrary computation, sensing, and communication.  Task-based models keep
progress at the granularity of tasks by repeatedly trying to re-execute each
task on power failure until it successfully finishes, only then moving on to
the next task.  Since these models do not rely on taking an expensive
checkpoint, it is usually faster than the checkpoint-based
solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
task-based programming model, where the programmer explicitly expresses the
application as a sequence of tasks and the transition between them.
%Task-based runtime systems ensure that repeated task executions are idempotent. 
%The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.  
%A programmer describes task control-flow as a {\em task graph}. Task flow happens at programmer demarcated transition points that may be conditional on program values. 
%Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.

@@

%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.  
% We do not need to emphasize the following figure! (Sinan)
%\begin{figure}
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
%	\caption{Cost of \emph{write} operation}
%\end{subfigure}%
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
%	\caption{Cost of \emph{read} operation}
%\end{subfigure}
%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
%\end{figure}

% As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
% the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
% voltage drop on a capacitor powering the device across 1600 read and write
% accesses to each memory, randomizing accesses to avoid caching effects. Our
% data showed that a non-volatile memory access consumes around 1.5 times the
% energy of a volatile access. The data confirm that, because non-volatile memory
% accesses are more expensive than volatile memory accesses, prior systems that
% frequently access non-volatile memory may be artificially inefficient.

%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. This observation is also consistent with the TI specification~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}

% \sys is motivated by this observation: \sys 
% moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
% accesses, to decrease energy consumption and increase the time between power
% failures.

@@

\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or
programmer-defined tasks can be both non-terminating and/or
inefficient.  If a task (or code between two checkpoints) consumes more than the fixed,
maximum energy that the device can buffer, then the task
will never be able to complete using buffered energy.  Such a task is
non-terminating, {\em prevents forward progress}, and makes the program
incorrect. 
%
If the task consumes far less energy than a device can buffer, the system
may operate {\em inefficiently}, saving the program state more often than needed.
%
Avoiding excessively costly, non-terminating, tasks and short, high-overhead,
tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}) are considered the problem of porting a program become much worse
because a large task on one device may be a relatively short task
on a device with a larger buffer.
%
\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which
merges multiple tasks to amortize the overhead when tasks are too small, and
{\em task downscaling}, which breaks up a task when the task is too large to complete. 
%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity. Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer. The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead. \sys's execution model based on {\em task coalescing} and {\em task downscaling}.  Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately. Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.
